---
title: "Model"
output: html_notebook
---

```{r}
library(data.table)
library(ggplot2)
```
Initialization of the individuals:

  - First, it has to be set the number of individuals that the model will have.
  - Each individual will be a vector of 4 components: the identity number, the 
    duplication of the DNA, the generation, and the biovolume.
  - The culture has also to be initialitzated, is a vector of the individuals.

```{r}
# the initial number of cells of the culture
num_cell = 4500

# initialization of the individuals variables
ident <- matrix(1:num_cell, 1, num_cell)
dupl <- matrix(0, 1, num_cell)
gen <- matrix(0, 1, num_cell)
biovol <- matrix(6000, 1, num_cell) + rnorm(num_cell, 0, 1000)

# culture generation as a vector of individuals
cultiu <- rbind(ident, biovol, dupl, gen)

# plot of the initial biovolume distribution
df <- data.frame(biovolum = cultiu[2,])
inici = as.numeric(df[,1])
hist(inici, breaks = 8, freq = FALSE, xlim = c(1000, 13000), xlab = "biovolume (µm^3)", ylab = "freq cells x 1.0e-5", main = "Initial distribution of Biovolumes", las = 1, yaxt='n')
axis(side=2, at=seq(0, 0.00035, 0.00005), labels=seq(0,35,5))
```
Now the model:

There will be 3 loops one inside the other. The most "exterior" one is the loop
over the days, for each day it will also be a loop over the hours, and at each
hour a loop over all the individuals of the culture.

For each individual at each hour the next processes are performed:
  - First, to check the reproduction conditions. If it fits the conditions then
    it will duplicate and a new cell will be added to the culture
  - Then each individual has to update its biovolume proportional to its surface

In the end, the logarithmic scale of the number of individuals per day is printed. 
Other results like the average biovolume per day or per hour or per day; the
evolution of one particular individual, the biovolume distribution at a concret
hour etc. can be also easily printed adding some lines to the code.
```{r}
# initialization of the days that the simulation will last
dia <- 1
dies_max <- 20

# creation of the vector with the number of individuals per day
n_indv <- dim(cultiu)[2]

# definition of gamma parameter
gam <- 0.05

# initialization of the vector with the boundary biovolumes
biovol_bound <- 8000
bio_dup <- matrix(biovol_bound, 1, dies_max) + rnorm(dies_max, 0, 300)

# starts the loop of the days
while (dia <= dies_max) {
  
  # initialization of the different gamma paramater per hour
  gama <- matrix(gam, 1, dies_max) + rnorm(12, 0, 0.001)
  
  # hours loop
  for (j in 1:12) {
    # each hour the cells will be visited in different order
    ordre <- sample(1:dim(cultiu)[2],dim(cultiu)[2],replace=F)
    
    # histogram of the biovolume saved at each hour
    df <- data.frame(biovolum = cultiu[2,])
    biovolum = as.numeric(df[,1])
    name = paste0('dia',paste0(dia, paste0('hora',paste0(j,".png"))))
    png(name)
    hist(biovolum, breaks = 8, freq = FALSE, xlim = c(1000, 13000), xlab = "biovolum (micrometres^3)", ylab = "freq cells") 
    abline(v = 8000, col="red", lwd=3, lty=2) 
    abline(v = 4000, col="red", lwd=3, lty=2)
    dev.off()
    
    # cells loop
    for (k in 1:dim(cultiu)[2]) {
      i <- ordre[k]
      
      # DUPLICATION
      # check if the cell biovolume is higher than the boundary one
      if (cultiu[2,i] >= bio_dup[dia]) {
        cultiu[3,i] = cultiu[3,i] + 1
      }
      
      # check if the cell fulfills the duplication conditions
      if (cultiu[3,i] == 2 & cultiu[4,i] <= 20) {
        # update of the old cell
        cultiu[2,i] = cultiu[2,i]/2 # the new biovolume
        cultiu[3,i] = 0
        cultiu[4,i] = cultiu[4,i] + 1 # new generation
        
        # initialization of the new cell
        cultiu <- cbind(cultiu, c(dim(cultiu)[2] + 1, cultiu[2,i], 0, cultiu[4,i]))
      }
      
      # INCOMES
      # calculation of the surface area
      superficie_i <- 4*pi*((3*cultiu[2,i])/(4*pi))^(2/3)
      
      # calculation of the new biovolume
      cultiu[2,i] = cultiu[2,i] + gama[dia]*superficie_i
    }
  }
  
  # day update
  dia = dia + 1
  # vector with the number of individuals at the end of each day
  n_indv <- append(n_indv, dim(cultiu)[2])
}

# graphic of log of the number of cells per day
plot(log10(n_indv), type="o", col="blue", axes = FALSE, ann = FALSE, ylim = range(log10(n_indv)))
box()
axis(1, las= 1)
axis(2, las= 1)
title(xlab="Time (days)")
title(ylab="ln(Number of cells)")
```